---
title: "Appendix A: Update to Fishery Data"
author: |
  | Tyler Jackson
  | Alaska Department of Fish and Game, tyler.jackson@alaska.gov
date: "`r format(Sys.time(), '%B %Y')`"
output:
  bookdown::pdf_document2:
    toc: no
header-includes:
   - \usepackage{float}
   - \usepackage{hanging}
   - \usepackage{enumerate}
   - \usepackage{longtable}
---
  
```{r, echo = F, message = F, warning = F}
library(gmacsr)
library(patchwork)
library(xtable)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.pos='H',
                      out.height = "\\textheight", out.width = "\\textwidth")

# load all out files
eag_may <- gmacs_read_allout("../../../AIGKC/models/2025/may/EAG/23.1c/Gmacsall.out", model_name = "eag_may", version = "2.20.21")
eag_sept <- gmacs_read_allout("../../../AIGKC/models/2025/sept/EAG/23.1c/Gmacsall.out", model_name = "eag_sept", version = "2.20.21")
wag_may <- gmacs_read_allout("../../../AIGKC/models/2025/may/WAG/23.1c/Gmacsall.out", model_name = "wag_may", version = "2.20.21")
wag_sept <- gmacs_read_allout("../../../AIGKC/models/2025/sept/WAG/23.1c/Gmacsall.out", model_name = "wag_sept", version = "2.20.21")

# get total catch time series
gmacs_get_catch_summary(list(eag_may, eag_sept, wag_may, wag_sept)) %>%
  filter(series == 2, year != 2024) %>%
  transmute(model, year, obs_catch) %>%
  pivot_wider(names_from = model, values_from = obs_catch) %>%
  transmute(year, eag_sept, eag_may, wag_sept, wag_may) -> totcatch_tab

# get total catch plot
gmacs_get_catch_summary(list(eag_may, eag_sept, wag_may, wag_sept)) %>%
  filter(series == 2, year != 2024) %>%
  separate(model, sep = "_", into = c("subdistrict", "data")) %>%
  mutate(subdistrict = toupper(subdistrict),
         data = ifelse(data == "sept", "Status Quo", "Update")) %>%
  ggplot()+
  geom_line(aes(x = factor(year), y = obs_catch, color = data, group = data))+
  facet_wrap(~subdistrict, ncol = 1)+
  scale_color_manual(values = cbpalette[c(1, 4)])+
  scale_y_continuous(labels = scales::comma)+
  scale_x_discrete(breaks = yraxis$breaks, labels = yraxis$labels)+
  labs(x = NULL, y = "Total Catch (t)", color = NULL) -> tot_plot

# plot of total size composition
gmacs_get_size_summary(list(eag_may, eag_sept)) %>%
  filter(mod_series == 2, year != 2024) %>%
  separate(model, sep = "_", into = c("subdistrict", "data")) %>%
  mutate(subdistrict = toupper(subdistrict),
         data = ifelse(data == "sept", "Status Quo", "Update")) %>%
  ggplot()+
  geom_area(aes(x = size, y = obs, fill = data), position = "identity", color = NA, alpha = 0.5)+
          scale_y_continuous(expand = expand_scale(mult = c(0, 0.1), add = c(0, 0)))+
          labs(x = "Carapace Length (mm)", y = NULL, color = NULL, fill = NULL, title = "EAG")+
          geom_text_npc(aes(npcx = "left", npcy = 0.6, label = year), check_overlap = T, size = 3)+
          facet_wrap(~year, ncol = 3, dir = "v")+
           scale_fill_manual(values = cbpalette[c(1, 4)])+
          theme(panel.spacing.x = unit(0.2, "lines"),
                panel.spacing.y = unit(0, "lines"),
                panel.border = element_blank(),
                axis.line.x = element_line(color = "grey70", size = 0.2),
                axis.ticks.y = element_blank(),
                axis.text.y = element_blank(),
                axis.text.x = element_text(size = 8),
                plot.title = element_text(hjust = 0.5),
                strip.background = element_blank(),
                strip.text.x = element_blank(),
                panel.background = element_blank()) -> eag_tot_comp_plot
gmacs_get_size_summary(list(wag_may, wag_sept)) %>%
  filter(mod_series == 2, year != 2024) %>%
  separate(model, sep = "_", into = c("subdistrict", "data")) %>%
  mutate(subdistrict = toupper(subdistrict),
         data = ifelse(data == "sept", "Status Quo", "Update")) %>%
  ggplot()+
  geom_area(aes(x = size, y = obs, fill = data), position = "identity", color = NA, alpha = 0.5)+
          scale_y_continuous(expand = expand_scale(mult = c(0, 0.1), add = c(0, 0)))+
          labs(x = "Carapace Length (mm)", y = NULL, color = NULL, fill = NULL, title = "WAG")+
          geom_text_npc(aes(npcx = "left", npcy = 0.6, label = year), check_overlap = T, size = 3)+
          facet_wrap(~year, ncol = 3, dir = "v")+
           scale_fill_manual(values = cbpalette[c(1, 4)])+
          theme(panel.spacing.x = unit(0.2, "lines"),
                panel.spacing.y = unit(0, "lines"),
                panel.border = element_blank(),
                axis.line.x = element_line(color = "grey70", size = 0.2),
                axis.ticks.y = element_blank(),
                axis.text.y = element_blank(),
                axis.text.x = element_text(size = 8),
                plot.title = element_text(hjust = 0.5),
                strip.background = element_blank(),
                strip.text.x = element_blank(),
                panel.background = element_blank()) -> wag_tot_comp_plot
```

# Purpose {-}

Ongoing efforts to standardized methods for summarizing fishery time series data resulted in minor changes to total catch and observer total size composition data for AIGKC. Difference are due to exclusion of pots with compramised biowtine and data filtering associated with estimation of average fishery weight (used in total catch estimation). Standardized methods for fishery data processing are detailed within the R package \textit{BSAIcrabR}, stored at \href{https://github.com/commfish/BSAIcrabR}{\textcolor{blue}{https://github.com/commfish/BSAIcrabR}}. Data processing specific to AIGKC are found here: \href{https://github.com/commfish/adfg_crab_observer/blob/master/aigkc/code/aigkc_BSAIcrabR.R}{\textcolor{blue}{aigkc\_BSAIcrabR.R}}. A comparison of time series data for which changes were made are detailed in table and figures below.


#### {-}

<!-- observed catch -->

\begin{table}[H]  
\centering
\caption{Comparison of estimates of total catch (t) between status quo and updated data summaries for the EAG and WAG. }

\label{table:totcatch}
\begin{tabular}{lcccccccccc}
& \multicolumn{2}{c}{EAG} &&  \multicolumn{2}{c}{WAG}\\
Year & Status Quo & Updated && Status Quo & Updated\\
\hline
```{r totcatch, results='asis', eval = T}

totcatch_tab %>%
  transmute(year = as.character(year), eag_sept, eag_may, space = NA, wag_sept, wag_may) -> x

# print table
print(xtable(x, align="lcccccc", digits =0),
			     only.contents=TRUE,
			     include.rownames=FALSE,
			     floating=FALSE,
			     include.colnames=FALSE,
			     hline.after=NULL,
			     format.args = list(big.mark = ","))

```
\hline
\end{tabular}
\end{table}

```{r totcatchp, fig.align='center', fig.cap="Comparison of estimates of total catch (t) between status quo and updated data summaries for the EAG and WAG. ", eval = T, fig.width = 6}
tot_plot
```

```{r eagtotcomp, fig.align='center', fig.cap="Comparison of estimates of total catch size composition between status quo and updated data summaries for the EAG. ", eval = T, fig.height = 7}
eag_tot_comp_plot
```

```{r wagtotcomp, fig.align='center', fig.cap="Comparison of estimates of total catch size composition between status quo and updated data summaries for the WAG. ", eval = T, fig.height = 7}
wag_tot_comp_plot
```