---
output:
  bookdown::pdf_document2:
    fig_caption: true
    includes: null
    toc: no
---
# Appendix C. Assessing uncertainty in model output due to lack of terminal year survey data for St. Matthew blue king crab {-}

\pagenumbering{gobble}

```{r global_options, include=FALSE, echo = FALSE, message = FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(fig.width = 12, fig.height = 7, echo = FALSE, warning = FALSE, message = FALSE,  dpi=300)
```

# Introduction{-}
The summer of 2020 trawl surveys were cancelled due to logistic difficulties caused by the global pandemic COVID-19. Therefore, the crab assessment authors meet to discuss approaches to address the potential of additional uncertainty in the current year models - specifically the projected mature male biomass and associated reference points. The objective of these approaches/simulations was to provide the plan team (CPT) and the SSC a range of potential additional uncertainty that could be applied to the buffers used on the OFL calculations to produce an appropriate ABC. 

## Objectives {-}

    1. Can we characterize the additional uncertainty in the current years estimates due to the lack of terminal year survey data? If so, what does it look like?

    2. Is the model uncertainty characterized in objective #1 currently included in the ABC buffer applied to this stock or do we need to apply additional uncertainty measures?


# Approaches {-}


## Approach 1 and 2: retrospective patterns with and without terminal survey data {-}

Retrospective analysis are typically performed on models to characterize the tendencies of a model to over or under estimate current trends in biomass, recruitment, etc. Retrospective patterns are described as a clear tendency for a model to either over or under estimate. Approach 1 compares the output of retrospective models with the terminal year of survey data and ones where the terminal year of trawl survey data are removed (both abundance and size composition data). 
A number of key model outputs were compared for these retrospective runs. These include: average recruitment, Bmsy, status of the stock, terminal year mmb, and reference point calculations (OFL). 

### Results {-}
In general models that lacked the terminal year of survey data performed similarly to models with the survey data. In cases where the model outputs differed the model without the terminal year of survey data tended to have results similar to the previous years model. 



## Approach 3: encompassing expected variability {-}

This approach was designed to run models with "fake" 2020 data to determine how much a data point in 2020 could have potential influenced the model outcome. The same key model outputs were compared in this approach as in approach 1. 

Copy methods here from Martin's write up. 

\pagebreak 

# Figures {-}
\newpage

```{r retrospective, fig.cap = "Retrospective run estimates of mature male biomass (mmb) for the SMBKC reference model (16.0) for the last 10 years. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/ssb_time_series_normal.png"), dpi = 125)
```


```{r retrospective2, fig.cap = "Retrospective run estimates of mature male biomass (mmb) for the SMBKC reference model (16.0) for the last 10 years, only showing the last 20 years for a detailed view. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/ssb_time_series_normal_current.png"), dpi = 125)
```

```{r retrospective-compare, fig.cap = "Retrospective run estimates of mature male biomass (mmb) for the SMBKC reference model (16.0) including models that eliminated the terminal year survey data. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/ssb_time_series_all.png"), dpi = 125)
```

```{r retrospective-compare2, fig.cap = "Retrospective run estimates of mature male biomass (mmb) for the SMBKC reference model (16.0) including models that eliminated the terminal year survey data for the last 5 model years. Highlighting the last 20 years for a more detailed view.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/ssb_time_series_last5_compare_recent.png"), dpi = 125)
```

```{r compare-avgR, fig.cap = "Comparison of average recruitment model estimates from 'normal' retrospective runs and those without the terminal year survey data.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/avgR_compare.png"), dpi = 125)
```

```{r compare-bmsy, fig.cap = "Comparison of Bmsy model estimates from 'normal' retrospective runs and those without the terminal year survey data.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/Bmsy_compare.png"), dpi = 125)
```

```{r compare-status, fig.cap = "Comparison of the model estimate of 'status' (B/Bmsy) from 'normal' retrospective runs and those without the terminal year survey data.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/Status_compare.png"), dpi = 125)
```

```{r compare-mmb, fig.cap = "Comparison of the model estimate of terminal year mmb from 'normal' retrospective runs and those without the terminal year survey data.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/Terminal_mmb_compare.png"), dpi = 125)
```


```{r compare-fofl, fig.cap = "Comparison of the model estimate of fofl from 'normal' retrospective runs and those without the terminal year survey data.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/fofl_compare.png"), dpi = 125)
```


```{r compare-OFL, fig.cap = "Comparison of the model estimate of OFL from 'normal' retrospective runs and those without the terminal year survey data.  "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/OFL_compare.png"), dpi = 125)
```


```{r app3-all-ssb, fig.cap = "Mature male biomass estimates from approach 3. Comparing the 2020 base model with a model that has a high 'fake' 2020 survey data point and one that has a low 'fake' survey data point. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/app3_ssb_all_yrs.png"), dpi = 125)
```

```{r app3-all-ssb2, fig.cap = "Mature male biomass estimates from approach 3. Comparing the 2020 base model with a model that has a high 'fake' 2020 survey data point and one that has a low 'fake' survey data point, only showing the last 20 years for detail on model differentiation. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/app3_last_10yrs_ssb.png"), dpi = 125)
```


```{r app3-compare, fig.cap = "Model output and reference points from approach 3. Comparing the 2020 base model with a model that has a high 'fake' 2020 survey data point and one that has a low 'fake' survey data point. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/app3_bar_graph_output.png"), dpi = 125)
```