---
output:
  bookdown::pdf_document2:
    fig_caption: true
    includes: null
    toc: no
---
# Appendix C. Explorations of survey coverage for St. Matthew blue king crab (SMBKC) {-}


```{r global_options, include=FALSE, echo = FALSE, message = FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(fig.width = 12, fig.height = 7, echo = FALSE, warning = FALSE, message = FALSE,  dpi=300)
library(xtable)
options(xtable.comment = FALSE)
```

\pagenumbering{arabic}
```{r page } 
#\pagenumbering{gobble}

```


# Introduction{-}
There are two index surveys included in the St. Matthew blue king crab (SMBKC) stock assessment model. The EBS NMFS Trawl survey has surveyed this stock annually (with the exception of 2020) since 1978 using the standard trawl grid around St. Matthew Island (Otto et al. 1978, XX). 
In 1995 ADF&G initiated a triennial pot survey of the area with the goal of surveying the same population and also to survey nearshore areas where females and juveniles were believed to be in higher numbers. The pot survey has 11 years of data, with more concentrated data in the mid-2010s when it was a annual survey from 2015 to 2018. There has not been a pot survey since 2018, but it is currently planned for fall of 2022. While both share the goal of providing an index of the SMBKC stock, the two surveys cover different geographic regions, and in some years they have encountered proportionally large numbers of male blue king crab in areas that are not covered by the other survey (map of survey areas).

In the last few years, the two surveys have presenting conflict trajectories for the spawning stock biomass (males) in the assessment model (Figure \@ref(fig:models2020)). Model runs in 2019 and 2020 attempting to explore this relationship further by looking at: model runs with only the NMFS trawl data (Figure \@ref(fig:models2020)), model runs with additional CVs (on both or one of the surveys), model runs with time blocks for q for the ADF&G pot survey, and just general model explorations of how these two surveys interact in the larger assessment model. 
In 2020, the year of the last full assessment, both the CPT and SSC suggested some explorations into the raw data from these surveys. The explorations presented here begin these explorations with the goal of determining the best path forward to advise on the assessment model structure.

## Spatial overlap {-}
Both surveys sample the area surrounding St.Matthew Island, however the spatial coverage and density of each differs. Figure \@ref(fig:map) visually displays the area of overlap between the two surveys (blue area), as well as the area occupied by each survey that does not overlap (grey and white). While the pot survey overlaps most of the area around St.Matthew's only the light blue area - the core 96 stations - are sampled in each year and therefore used in the assessment. The north east coast of the island contains the NMFS station R-24 which was been a "hot spot" in past survey years. However, the ADF&G survey data from this area is inconsistent and therefore not used in the assessment. 
In addition

## Pot survey {-}
The ADF&G pot survey was also initiated to survey areas around St.Matthew Island that have rocky bottom conditions where legal male and mature female crab are at high densities (Pengilly and Vanek 2014). 

## Objectives {-}

    1. Explore the spatial difference between the NMFS trawl survey and the ADF&G pot survey for SMBKC

        a. How does the mature male biomass compare when using the same spatial footprint?
        b. What is the contribution of R-24 to the NMFS trawl survey data set?
        c. What are the potential model parameterization options?


# Approaches and Findings {-}



# Results {-}
Retrospective analysis of the base model show a retrospective pattern that tends to overestimate mature male biomass (MMB) in the terminal year (Figure \@ref(fig:retrospective) and \@ref(fig:retrospective2)). Using a peel of the last 5 years estimates of MMB the estimated Mohn's $\rho$ is -0.346, which suggests a retrospective pattern in the MMB estimates for the base model. Since 2018 the MMB estimates have been relatively stable, however, they are the lowest in the model history and reflect a time of overfished declaration for the stock. 



# Recommendations  {-}




\pagebreak 

# Figures {-}

```{r map, fig.cap = "Map to visualize the spatial coverage and overlap of the two surveys. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/figures/pot_and_trawl_stations_map.png"), dpi = 125)
```


```{r models2020, fig.cap = "Model scenarios from 2020 SAFE document, which included a *no pot* exploration. "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_22/doc/safe_figure/2020mod_scen_ssb_wprojected_yr.png"), dpi = 125)
```

```{r cpue, fig.cap = "Survey data for both the ADF&G pot survey (CPUE) and the NMFS trawl survey (design based area swept). "}
knitr::include_graphics(paste0(here::here(), "/SMBKC/smbkc_22/doc/safe_figure/cpue.png"), dpi = 125)
```



\pagebreak 
# Tables {-}

```{r diff-compare, results = "asis"}
df <- read.csv(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/stats_summary_compare_table2.csv"))

names(df) <- c("Year","AvgR","Bmsy", "Terminal MMB", "Status", "Fofl", "OFL")
tab <- xtable(df, caption = "Comparisons of the percent difference in parameter estimates for the retrospective models with and without the terminal year of survey data.", label = "tab:diff-compare", digits = 3)
print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x}, NA.string = "-")
```

```{r diff-cv, results = "asis"}
df <- read.csv(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/CV_summary_compare_table.csv"))

names(df) <- c("Type","CV-Bmsy","CV-OFL", "CV-status", "CV-terminal-SSB")
tab <- xtable(df, caption = "Average CV over all years (2010-2019) for normal retrospective runs and those missing the terminal year of survey data. ", label = "tab:diff-cv", digits = 2)
print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x}, NA.string = "-")
```


```{r app3-results, results = "asis"}
df <- read.csv(paste0(here::here(), "/SMBKC/smbkc_20/retrospective_model_1/figures/stats_summary_app3_table.csv"))

names(df) <- c("Variable","Diff-Ltobase","Diff-Htobase")
tab <- xtable(df, caption = "Comparisons of the percent difference in parameter estimates for the low and high models in approach 3 compared to the 2020 base model (16.0).", label = "tab:app3-results", digits = 3)
print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x}, NA.string = "-")
```