---
title: "Saint Matthew Island Blue King Crab Stock Assesssment 2019"
author: |
  | Katie Palof$^1$, Jie Zheng$^1$, Jim Ianelli$^2$ 
  | $^1$Alaska Department of Fish and Game, katie.palof@alaska.gov and jie.zheng@alaska.gov
  | $^2$NOAA, jim.ianelli@noaa.gov
date: "`r format(Sys.time(), '%B %Y')`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    includes: null
    toc: no
---

```{r global_options, include=FALSE, echo = FALSE, message = FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,  dpi=300)
```

```{r, load_packages, echo = FALSE, include = FALSE}
library(gmr)
library(xtable)
options(xtable.comment = FALSE)

# the model specs
mod_names <- c("model 18.0", "model 19.0 (reference)", "model 19.1 (survey fit)") 
.MODELDIR = c(paste0(here::here(), "/SMBKC/smbkc_18a/model_1/"), paste0(here::here(), "/SMBKC/smbkc_19/model_1/initial_run/"),
              paste0(here::here(), "/SMBKC/smbkc_19/model_5/")) #need to update these model options
.THEME    = theme_bw(base_size = 12, base_family = "")
.OVERLAY  = TRUE
.SEX      = c("Aggregate","Male")
.FLEET    = c("Pot","Trawl bycatch","Fixed bycatch","NMFS Trawl","ADF&G Pot")
.TYPE     = c("Retained & Discarded","Retained","Discarded")
.SHELL    = c("Aggregate")
.MATURITY = c("Aggregate")
.SEAS     = c("1","2","3","4","5")
# Read report file and create gmacs report object (a list):
fn       <- paste0(.MODELDIR, "gmacs")
M        <- lapply(fn, read_admb)
names(M) <- mod_names
jj <- 1 # The position in the list that Jies model outputs sit

nmult_1 <- 1e+06
nmult_2 <- 0.0004535923 * 1e+6
#fn <- paste0(.MODELDIR[1], "gmacs")
#Mmatch <- lapply(fn, read_admb)
#names(Mmatch) <- c("SMBKC")

fn <- paste0(.MODELDIR[2], "gmacs")
Mbase <- lapply(fn, read_admb)
names(Mbase) <- c("SMBKC")

rinline <- function(code){
  html <- '<code  class="r">``` `CODE` ```</code>'
  sub("CODE", code, html)
}

ref_mod <- 2 # base
rec_mod <- 2 # base
mod_scen<- 2:3 #scenarios you want graphed together
```

```{r exesum, echo = FALSE}
.get_cpue_df(Mbase) %>% 
  filter(fleet==.FLEET[4]) %>% 
  mutate(x = round(100*pred/mean(pred),0)) %>% 
  select(x) %>% 
  tail(1) %>%
  .$x -> bio_lt_percent
```

\pagenumbering{gobble}

# Executive Summary {-}
1. **Stock**: Blue king crab, *Paralithodes platypus*, Saint Matthew Island (SMBKC), Alaska.

2. **Catches**: Peak historical harvest was 4,288 t (9.454 million pounds) in 1983/84^[1983/84 refers to a fishing
year that extends from 1 July 1983 to 30 June 1984.]. 
The fishery was closed for 10 years after the stock was declared overfished in 1999. 
Fishing resumed in 2009/10 with a fishery-reported retained catch of 209 t (0.461 million
pounds), less than half the 529.3 t (1.167 million pound) TAC.
Following three more years of modest harvests
supported by a fishery catch per unit effort (CPUE) of around 10 crab per pot lift, the fishery was again closed in
2013/14 due to declining trawl-survey estimates of abundance and concerns about the health of the stock. 
The directed fishery resumed again in 2014/15 with a TAC of 300 t (0.655 million pounds), but the fishery performance was
relatively poor with a retained catch of 140 t (0.309 million pounds). 
The retained catch in 2015/16 was even lower at 48 t (0.105 million pounds) and the fishery has remained closed since 2016/17.

3. **Stock biomass**: The 1978-2018 NMFS trawl survey mean biomass is 5,605 t with the 2019 value being the
15th lowest (3,170 t; the tenth lowest since 2000). 
This 2019 biomass of $\ge$ 90 mm carapace length (CL) male crab is 
57% of the long term mean  at 6.99 million pounds (with a CV of 34%), and is a 83% increase from the 2018 mean biomass.
The most recent 3-year average of the NMFS survey is 40% of the mean value, indicating a decline in biomass compared to
historical survey estimates, notably in 2010 and 2011 that were over four times the current average.
However, the 2019 value is substantially larger than the two previous years (3,170 t compared to 1,731 t in 2018 and 1,794 t in 2017).
The ADFG pot survey was did not take place in 2019, but in 2018 the relative biomass in this index was the lowest
in the time series (12% of the mean from the 11 surveys conducted since 1995).
The assessment model estimates tempers this increase and suggests that the stock (in survey biomass units) is presently at about
`r bio_lt_percent `% 
of the long term model-predicted survey biomass average. 
The trend from these values suggests a substantial increase, which does not fit the observed survey data points well.

4. **Recruitment**: Recruitment is based on estimated number of male crab within the 90-104 mm CL size class
in each year. The 2019 trawl-survey area-swept estimate of 0.403 million male SMBKC in this size class is the twelfth lowest in 
the 42 years since 1978 and follows two of the lowest previously observed values in 2017 and 2018. The recent six-year (2014 - 2019) average recruitment 
is only 47% of this mean. 
In the pot-survey, the abundance of this size group in 2017 was also the second-lowest in the time series (22% of the mean for the 
available pot-survey data) whereas in 2018 the value was the lowest observed at only 10% of the mean value.

5. **Management performance**: In this assessment estimated total male catch is the sum of fishery-reported retained
catch, estimated male discard mortality in the directed fishery, and estimated male bycatch mortality in the groundfish
fisheries.
Based on the reference model for SMBKC, the estimate for mature male biomass is below the minimum stock-size 
threshold (MSST) in 2017/18 and is hence is in an "overfished" condition, despite fishery closures in the last two years 
(and hence overfishing has not occurred) (Tables ). 
Computations which indicate the relative impact of fishing 
(i.e., the "dynamic $B_0$") suggests that the current spawning stock biomass has been reduced to 60%
<!--- `r #.get_dynB0_df(Mbase) %>% transmute(round(100*ssb,0)) %>% slice(n())`%  -->
 of what it would have been in the absence of fishing, assuming the same level of recruitment as estimated. 
 
 
