---
title: "Saint Matthew Island Blue King Crab Stock Assessment 2018a"
author: "| Katie Palof$^1$, Jie Zheng$^1$, Jim Ianelli$^2$ \n|
  $^1$Alaska Department of Fish and Game, katie.palof@alaska.gov and jie.zheng@alaska.gov
  \n| $^2$NOAA, jim.ianelli@noaa.gov\n|"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  pdf_document:
    highlight: zenburn
    includes: null
    toc: no
  word_document:
    toc: yes
---

```{r global_options, include=FALSE }
library(knitr)
opts_chunk$set(fig.width = 12, fig.height = 7, echo = FALSE, warning = FALSE, message = FALSE,  dpi=300)
#opts_chunk$set(fig.width = 12, fig.height = 7, echo = FALSE, warning = FALSE, message = FALSE,  dpi=72)
```

```{r, load_packages, include=FALSE}
library(gmr)
library(xtable)
options(xtable.comment = FALSE)

# The model specs
#mod_names <- c("Andre","2018 BTS","Reference","VAST","Fit surveys") #,"Gmacs M","Gmacs Francis","fit surveys")
#mod_names <- c("Base Model 1", "Base Model 1a", "VAST", "VAST a", "Fit surveys", "Fit surveys a") #,"2018 BTS","Reference","VAST","Fit surveys")
mod_names <- c("Base Model 1", "VAST", "Fit surveys") 
.MODELDIR = c(paste0(here::here(), "/SMBKC/smbkc_18a/model_1/"), paste0(here::here(), "/SMBKC/smbkc_18a/model_4/"), paste0(here::here(), "/SMBKC/smbkc_18a/model_5/"))
.THEME    = theme_bw(base_size = 12, base_family = "")
.OVERLAY  = TRUE
.SEX      = c("Aggregate","Male")
.FLEET    = c("Pot","Trawl bycatch","Fixed bycatch","NMFS Trawl","ADF&G Pot")
.TYPE     = c("Retained & Discarded","Retained","Discarded")
.SHELL    = c("Aggregate")
.MATURITY = c("Aggregate")
.SEAS     = c("1","2","3","4","5")

# Read report file and create gmacs report object (a list):
fn       <- paste0(.MODELDIR, "gmacs")
M        <- lapply(fn, read_admb)
names(M) <- mod_names

nmult_1 <- 1e+06
nmult_2 <- 0.0004535923 * 1e+6
#fn <- paste0(.MODELDIR[1], "gmacs")
#Mmatch <- lapply(fn, read_admb)
#names(Mmatch) <- c("SMBKC")

fn <- paste0(.MODELDIR[1], "gmacs")
Mbase <- lapply(fn, read_admb)
names(Mbase) <- c("SMBKC")

rinline <- function(code){
  html <- '<code  class="r">``` `CODE` ```</code>'
  sub("CODE", code, html)
}

ref_mod <- 1 # base
rec_mod <- 1 # base
mod_scen<- 3:5
```


# 2019 Executive Summary 
1. **Stock**: Blue king crab, *Paralithodes platypus*, Saint Matthew Island (SMBKC), Alaska.

2. **Catches**: Peak historical harvest was 4,288 t (9.454 million pounds) in 1983/84^[1983/84 refers to a fishing
year that extends from 1 July 1983 to 30 June 1984.]. The fishery was closed for 10 years after the stock was declared
overfished in 1999. Fishing resumed in 2009/10 with a fishery-reported retained catch of 209 t (0.461 million
pounds), less than half the 529.3 t (1.167 million pound) TAC. Following three more years of modest harvests
supported by a fishery catch per unit effort (CPUE) of around 10 crab per pot lift, the fishery was again closed in
2013/14 due to declining trawl-survey estimates of abundance and concerns about the health of the stock. 
The directed fishery resumed again in 2014/15 with a TAC of 300 t (0.655 million pounds), but the fishery performance was
relatively poor with a retained catch of 140 t (0.309 million pounds). 
The retained catch in 2015/16 was even lower at 48 t (0.105 million pounds) and the fishery has remained closed since 2016/17.

3. **Stock biomass**: The 1978-2018 NMFS trawl survey mean biomass is 5,605 t with the 2019 value being the
15th lowest (3,170 t; the tenth lowest since 2000). This 2019 biomass of $\ge$ 90 mm carapace length (CL) male crab is 
57% of the long term mean  at 6.99 million pounds (with a CV of 34%), and is a 83% increase from the 2018 mean biomass.

The most recent 3-year average of the NMFS survey is 40% of the mean value, indicating a decline in biomass compared to
historical survey estimates, notably in 2010 and 2011 that were over four times the current average. However, the 2019 value is substantially larger than the two previous years (3,170 t compared to 1,731 t in 2018 and 1,794 t in 2017).
The ADFG pot survey was did not take place in 2019, but in 2018 the relative biomass in this index was the lowest
in the time series (12% of the mean from the 11 surveys conducted since 1995).
The assessment model estimates appears exagerate this increase and suggests that the stock (in survey biomass units) is presently at about
`r .get_cpue_df(Mbase) %>% filter(fleet==.FLEET[4]) %>% transmute(x = round(100*pred/mean(pred),0)) %>% tail(1) %>% .$x `% 
of the long term model-predicted survey biomass average. The trend from these values suggests a substantial increase, which does not fit the observed survey data points well.

4. **Recruitment**: Recruitment is based on estimated number of male crab within the 90-104 mm CL size class
in each year. The 2019 trawl-survey area-swept estimate of 0.403 million male SMBKC in this size class is the twelfth lowest in 
the 42 years since 1978 and follows two of the lowest previously observed values in 2017 and 2018. The recent six-year (2014 - 2019) average recruitment 
is only 47% of this mean. 
In the pot-survey, the abundance of this size group in 2017 was also the second-lowest in the time series (22% of the mean for the 
available pot-survey data) whereas in 2018 the value was the lowest observed at only 10% of the mean value.

5. **Management performance**: In this assessment estimated total male catch is the sum of fishery-reported retained
catch, estimated male discard mortality in the directed fishery, and estimated male bycatch mortality in the groundfish
fisheries.  Based on the reference model for SMBKC, the estimate for mature male biomass is below the minimum stock-size 
threshold (MSST) in 2017/18 and is hence is in an "overfished" condition, despite fishery closures in the last two years 
(and hence overfishing has not occurred) (Tables \ref{tab:status} and \ref{tab:status_pounds}). 
Computations which indicate the relative impact of fishing 
(i.e., the "dynamic $B_0$") suggests that the current spawning stock biomass has been reduced to 60%
<!--- `r #.get_dynB0_df(Mbase) %>% transmute(round(100*ssb,0)) %>% slice(n())`%  -->
 of what it would have been in the absence of
fishing, assuming the same level of recruitment as estimated. 

\begin{table}[ht]
\centering
\caption{Status and catch specifications (1000 t) for the reference model.
A - calculated from the assessment reviewed by the Crab Plan Team in September 2014,    
B - calculated from the assessment reviewed by the Crab Plan Team in September 2015,    
C - calculated from the assessment reviewed by the Crab Plan Team in September 2016,   
D - calculated from the assessment reviewed by the Crab Plan Team in September 2017,   
E - calculated from the 2018 model approved by the Crab Plan Team in September 2018 with GMACS modifications. 
}
\label{tab:status}
\begin{tabular}{lccccccc}
  \hline
  & & Biomass & & Retained & Total & & \\ 
  Year & MSST & ($\mathit{MMB}_\text{mating}$) & TAC & catch & male catch & OFL & ABC \\ 
  \hline
  2013/14 & 1.50$^A$ & 3.01$^A$ & 0.00 & 0.00 & 0.00 & 0.56 & 0.45 \\ 
  2014/15 & 1.86$^B$ & 2.48$^B$ & 0.30 & 0.14 & 0.15 & 0.43 & 0.34 \\ 
  2015/16 & 1.84$^C$ & 2.11$^C$ & 0.19 & 0.05 & 0.053 & 0.28 & 0.22 \\ 
  2016/17 & 1.97$^D$ & 2.23$^D$ & 0.00 & 0.00 & 0.001 & 0.14 & 0.11 \\ 
  2017/18 & 1.85     & 1.29     & 0.00 & 0.00 & 0.003 & 0.12 & 0.10
  2018/19 & `r round(M[[rec_mod]]$spr_bmsy/1000 * 0.5, 2)`$^E$ & `r round(M[[rec_mod]]$ssb[length(M[[rec_mod]]$ssb)]/1000, 2)`$^E$ & 0.00 & 0.00 & 0.001 & 0.04 & 0.03 \\
  2019/20 & & `r round(M[[rec_mod]]$spr_bmsy*M[[rec_mod]]$spr_depl/1000, 2)`$^E$ & & & & `r round(M[[rec_mod]]$spr_cofl/1000, 2)` & `r round(M[[rec_mod]]$spr_cofl/1000*0.8, 2)` \\
 \\
  \hline
\end{tabular}
\end{table}
  <!-- \multicolumn{8}{p{\linewidth}}{l}{ -->
\begin{table}[ht]
\centering
\caption{Status and catch specifications (million pounds) for the reference model.} 
\label{tab:status_pounds}
\begin{tabular}{lccccccc}
  \hline
  & & Biomass & & Retained & Total & & \\ 
  Year & MSST & ($\mathit{MMB}_\text{mating}$) & TAC & catch & male catch & OFL & ABC \\ 
  \hline
  2013/14 & 3.4$^A$ & 6.64$^A$ & 0.000 & 0.000 & 0.001  & 1.24 & 0.99 \\ 
  2014/15 & 4.1$^B$ & 5.47$^B$ & 0.655 & 0.309 & 0.332  & 0.94 & 0.75 \\ 
  2015/16 & 4.1$^C$ & 4.65$^C$ & 0.419 & 0.110 & 0.117  & 0.62 & 0.49 \\ 
  2016/17 & 4.3$^D$ & 4.91$^D$ & 0.000 & 0.000 & 0.002  & 0.31 & 0.25 \\ 
  2017/18 & 4.1$^E$ & 2.85$^E$ & 0.000 & 0.000 & 0.002  & 0.31 & 0.25 \\ 
  2017/18 & `r sprintf("%.1f", round(M[[rec_mod]]$spr_bmsy * 2204.62/1e6 * 0.5, 1))`$^E$ & `r round(M[[rec_mod]]$ssb[length(M[[rec_mod]]$ssb)-1] * 2204.62/1e6, 2)`$^E$ & 0.000 & 0.000 & 0.007 & 0.27 & 0.22 \\
  2018/19 & & `r round(M[[rec_mod]]$ssb[length(M[[rec_mod]]$ssb)] * 2204.62/1e6, 2)`$^E$ & & & & `r round(M[[rec_mod]]$spr_cofl * 2204.62/1e6, 2)` & `r round(M[[rec_mod]]$spr_cofl * 2204.62/1e6*0.8, 2)` \\
  2019/20 \\
  \hline
\end{tabular}
\end{table}



6. **Basis for the OFL**: Estimated mature-male biomass (MMB) on 15 February is used as the measure of 
 biomass for this Tier 4 stock, with males measuring $\ge$ 105 mm CL 
 considered mature. The $B_\mathit{MSY}$ proxy is obtained by averaging estimated MMB over a specific reference period, and current CPT/SSC guidance recommends using the full assessment time frame as the default reference period (Table \ref{tab:ofl_basis}).
\begin{table}[ht]
\centering
\caption{Basis for the OFL (1000 t) from the reference model.} 
\label{tab:ofl_basis}
\begin{tabular}{lcccccccc}
  \hline
  & & & Biomass & & & & & Natural \\ 
  Year & Tier & $B_\mathit{MSY}$ & ($\mathit{MMB}_\text{mating}$) & $B/B_\mathit{MSY}$ & $F_\mathit{OFL}$ & $\gamma$ & Basis for $B_\mathit{MSY}$ & mortality \\ 
  \hline
  2013/14 & 4b & 3.06 & 3.01 & 0.98 & 0.18 & 1 & 1978-2013 & 0.18 \\ 
  2014/15 & 4b & 3.28 & 2.71 & 0.82 & 0.14 & 1 & 1978-2014 & 0.18 \\ 
  2015/16 & 4b & 3.71 & 2.45 & 0.66 & 0.11 & 1 & 1978-2015 & 0.18 \\ 
  2016/17 & 4b & 3.67 & 2.23 & 0.61 & 0.09 & 1 & 1978-2016 & 0.18 \\ 
  2017/18 & 4b & 3.86 & 2.05 & 0.53 & 0.08 & 1 & 1978-2017 & 0.18 \\ 
  2018/19 & 4b & `r round(M[[rec_mod]]$spr_bmsy/1000, 2)` & 
  `r round(M[[rec_mod]]$ssb[length(M[[rec_mod]]$ssb)]/1000, 2)` & 
  `r round(M[[rec_mod]]$ssb[length(M[[rec_mod]]$ssb)]/M[[rec_mod]]$spr_bmsy, 2)` & 
  `r round(M[[rec_mod]]$sd_fofl[1], 3) ` & 
  1 & 1978-2018 & 0.18 \\ 
  2019/20 \\
  \hline
\end{tabular}
\end{table}

# A. Summary of Major Changes

## Changes in Management of the Fishery

There are no new changes in management of the fishery.

## Changes to the Input Data

Data used in this assessment have been updated to include the most recently
available fishery and survey numbers. This assessment makes use of one new
survey data point - the 2019 NMFS trawl-survey estimate of abudance but no ADF&G pot survey was conducted in 2019. Both of these surveys have associated size
compositon data. The assessment also uses updated 2010-2018 groundfish and
fixed gear bycatch estimates based on NMFS Alaska Regional Office (AKRO) data. 
The directed fishery has been closed since 2016/17 so fishery data in recent years are unavailable.

## Changes in Assessment Methodology

This assessment uses the General model for Alasks crab stocks (Gmacs) framework. The
model is configured to track three stages of length categories and was 
first presented in May 2011 by Bill Gaeuman and accepted by the CPT in May 2012. 
A difference from the original approach, and that used here, is that natural and fishing mortality are
continuous within `r M[[1]]$nseason` discrete seasons (using the appropriate
catch equation rather than assuming an applied pulse removal). Season length in Gmacs
is controlled by changing the proportion of natural mortality that is applied
each season. Diagnostic output includes estimates of the "dynamic $B_0$" which simply
computes the ratio of the spawning biomass as estimated relative to the spawning biomass that would have occurred 
had there been no historical fishing mortality. 
Details of this implementation and other model details are provided in Appendix A (re-order appendices).

## Changes in Assessment Results

Both surveys indicate a decline over the past few years. The "reference" model is that which was selected for use in 2018.  There is only one new data set to be included this year so this becomes the updated reference model. Two alternative models are presented for
sensitivity. One involves a re-analysis of the NMFS trawl survey data using a spatio-temporal Delta-GLMM approach (VAST
model; Thorson and Barnett 2017) and the other configuration (named "Fit survey") simply adds emphasis on 
the design-based survey data
(by assuming a lower input variance). The VAST model suggests a XXX from the 2018 survey estimate. However,
 the model tends to moderate the noise in the survey observations and declines.
 
# B. Responses to SSC and CPT Comments

## CPT and SSC Comments on Assessments in General
Not updated since 2018

Comment: *Regarding general code development, the SSC and CPT outstanding requests continue to be as follows:*

  1. *add the ability to conduct retrospective analyses*

     Progress was limited in implementing this feature.

  1. *add ability to estimate bycatch fishing mortality rates when observer data are missing but effort data is available*

     This was completed.

  1. *Continued exploration of data weighting (Francis and other approaches) and evaluation of models with and without the 
     1998 natural mortality spike. The authors are encouraged to bring other models forward for CPT and SSC consideration*

      We continued to include an alternative time series estimated from the NMFS trawl survey using the VAST spatiotemporal Delta GLMM model 
      and continued with the iterative re-weighting for composition data.

# C. Introduction

## Scientific Name

The blue king crab is a lithodid crab, *Paralithodes platypus* (Brant 1850).

## Distribution

Blue king crab are sporadically distributed throughout the North Pacific Ocean from Hokkaido, Japan, to southeastern
Alaska (Figure ).  In the eastern Bering Sea small populations are distributed around St. Matthew
Island, the Pribilof Islands, St. Lawrence Island, and Nunivak Island. Isolated populations also exist in some other
cold water areas of the Gulf of Alaska (NPFMC 1998). The St. Matthew Island Section for blue king crab is within Area Q2
(Figure ), which is the Northern District of the Bering Sea king crab registration area and
includes the waters north of Cape Newenham (58&deg;39' N. lat.) and south of Cape Romanzof (61&deg;49' N. lat.).

## Stock Structure

The Alaska Department of Fish and Game (ADF&G) Gene Conservation Laboratory, has detected regional population
differences between blue king crab collected from St. Matthew Island and the Pribilof Islands^[NOAA grant Bering Sea
Crab Research II, NA16FN2621, 1997.]. NMFS tag-return data from studies on blue king crab in the Pribilof Islands and
St. Matthew Island support the idea that legal-sized males do not migrate between the two areas (Otto and Cummiskey
1990). St. Matthew Island blue king crab tend to be smaller than their Pribilof conspecifics, and the two stocks are
managed separately.

## Life History

Like the red king crab, *Paralithodes camtshaticus*, the blue king crab is considered a shallow water species by
comparison with other lithodids such as golden king crab, *Lithodes aequispinus*, and the scarlet king crab, 
*Lithodes couesi* (Donaldson and Byersdorfer 2005). Adult male blue king crab are found at an average depth of 70 m (NPFMC 1998).
The reproductive cycle appears to be annual for the first two reproductive cycles and biennial thereafter (Jensen
and Armstrong 1989), and mature crab seasonally migrate inshore where they molt and mate. Unlike red king crab, juvenile
blue king crab do not form pods, but instead rely on cryptic coloration for protection from predators and require
suitable habitat such as cobble and shell hash. Somerton and MacIntosh (1983) estimated SMBKC male size at sexual
maturity to be 77 mm carapace length (CL). Paul et al. (1991) found that spermatophores were present in the vas deferens
of 50% of the St. Matthew Island blue king crab males examined with sizes of 40-49 mm CL and in 100% of the males at
least 100 mm CL. Spermataphore diameter also increased with increasing CL with an asymptote at ~ 100 mm CL. It was noted,
however, that although spermataphore presence indicates physiological sexual maturity, it may not be an indicator of
functional sexual maturity. For purposes of management of the St. Matthew Island blue king crab fishery, the State of
Alaska uses 105 mm CL to define the lower size bound of functionally mature males (Pengilly and Schmidt 1995). Otto and
Cummiskey (1990) report an average growth increment of 14.1 mm CL for adult SMBKC males.

## Management History

The SMBKC fishery developed subsequent to baseline ecological studies associated with oil exploration (Otto 1990). Ten
U.S. vessels harvested 545 t (1.202 million pounds) in 1977, and harvests peaked in 1983 when 164 vessels landed 4,288 t
(9.454 million pounds) (Fitch et al. 2012; Table \ref{tab:smbkc_fishery}).

The fishing seasons were generally short, often lasting only a few days. The fishery was declared overfished and closed
in 1999 when the stock biomass estimate was below the minimum stock-size threshold (MSST) of 4,990 t (11.0 million
pounds) as defined by the Fishery Management Plan (FMP) for the Bering Sea/Aleutian Islands King and Tanner crabs (NPFMC
1999). Zheng and Kruse (2002) hypothesized a high level of SMBKC natural mortality from 1998 to 1999 as an explanation
for the low catch per unit effort (CPUE) in the 1998/99 commercial fishery and the low numbers across all male crab size
groups caught in the annual NMFS eastern Bering Sea trawl survey from 1999 to 2005 (see survey data in next section). In
November 2000, Amendment 15 to the FMP for Bering Sea/Aleutian Islands king and Tanner crabs was approved to implement a
rebuilding plan for the SMBKC stock (NPFMC 2000). The rebuilding plan included a State of Alaska regulatory harvest strategy 
(*5 AAC 34.917*), area closures, and gear modifications. In addition, commercial crab fisheries near St. Matthew Island were
scheduled in fall and early winter to reduce the potential for bycatch mortality of vulnerable molting and mating crab.

NMFS declared the stock rebuilt on 21 September 2009, and the fishery was reopened after a 10-year closure on 15 October
2009 with a TAC of 529 t (1.167 million pounds), closing again by regulation on 1 February 2010. Seven participating
vessels landed a catch of 209 t (0.461 million pounds) with a reported effort of 10,697 pot lifts and an estimated CPUE of
9.9 retained individual crab per pot lift. The fishery remained open the next three years with modest harvests and
similar CPUE, but large declines in the NMFS trawl-survey estimate of stock abundance raised concerns about the health
of the stock. This prompted ADF&G to close the fishery again for the 2013/14 season. 
The fishery was reopened for the 2014/15 season with a low TAC of 297 t (0.655 million pounds) and in 2015/16 the TAC
was further reduced to 186 t (0.411 million pounds) then completely closed during the 2016/17 season.

Although historical observer data are limited due to low sampling effort, bycatch of female and sublegal male crab from
the directed blue king crab fishery off St. Matthew Island was relatively high historically, with estimated total
bycatch in terms of number of crab captured sometimes more than twice as high as the catch of legal crab (Moore et al.
2000; ADF&G Crab Observer Database). Pot-lift sampling by ADF&G crab observers (Gaeuman 2013; ADF&G Crab Observer
Database) indicates similar bycatch rates of discarded male crab since the reopening of the fishery (Table
\ref{tab:stage_cpue_1}), with total male discard mortality in the 2012/13 directed fishery estimated at about 12% (88
t or 0.193 million pounds) of the reported retained catch weight, assuming 20% handling mortality.

These data suggest a reduction in the bycatch of females, which may be attributable to the later timing of the
contemporary fishery and the more offshore distribution of fishery effort since reopening in 2009/10^[D. Pengilly,
ADF&G, pers. comm.]. Some bycatch of discarded blue king crab has also been observed historically in the eastern Bering
Sea snow crab fishery, but in recent years it has generally been negligible.  The St. Matthew Island golden king crab
fishery, the third commercial crab fishery to have taken place in the area,  typically occurred in areas with depths
exceeding blue king crab distribution. The NMFS observer data suggest that variable, but mostly limited, SMBKC bycatch has
also occurred in the eastern Bering Sea groundfish fisheries (Table \ref{tab:smbkc_groundfish_bycatch}).

# D. Data

## Summary of New Information

Data used in this assessment were updated to include the most recently available fishery and survey numbers. This
assessment makes use of two new survey data points including the 2018 NMFS trawl-survey estimate of abudance, and the
2018 ADF&G pot survey CPUE. Both of these surveys have associated size compositon data. The assessment also uses updated
1993-2016 groundfish and fixed gear bycatch estimates based on AKRO data. The fishery was closed in 2016/17 so no
directed fishery catch data were available. The data used in each of the new models is shown in Figure
\ref{fig:data_extent}.

## Major Data Sources

Major data sources used in this assessment include annual directed-fishery retained-catch statistics from fish tickets
(1978/79-1998/99, 2009/10-2012/13, and 2014/15-2015/16; Table \ref{tab:smbkc_fishery}); results from the annual NMFS
eastern Bering Sea trawl survey (1978-2018; Table \ref{tab:stage_cpue_nmfs}); results from the ADF&G SMBKC pot
survey (every third year during 1995-2013, then 2015-2018; Table \ref{tab:stage_cpue});
mean somatic mass given length category by year (Table \ref{tab:length_weight}); 
size-frequency information from ADF&G crab-observer pot-lift sampling (1990/91-1998/99, 2009/10-2012/13, and
2014/15-2016/17; Table \ref{tab:stage_cpue_1}); and the NMFS groundfish-observer bycatch biomass estimates (1992/93-2016/17;
Table \ref{tab:smbkc_groundfish_bycatch}).

Figure \ref{fig:stations} maps stations from which SMBKC trawl-survey and pot-survey data were obtained. Further
information concerning the NMFS trawl survey as it relates to commercial crab species is available in Daly et al.
(2014); see Gish et al. (2012) for a description of ADF&G SMBKC pot-survey methods. It should be noted that the two
surveys cover different geographic regions and that each has in some years encountered proportionally large numbers of
male blue king crab in areas not covered by the other survey (Figure \ref{fig:catch181}). Crab-observer sampling
protocols are detailed in the crab-observer training manual (ADF&G 2013). Groundfish SMBKC bycatch data come from the NMFS
Regional office and have been compiled to coincide with the 
SMBKC management area.


## Other Data Sources

The growth transition matrix used is based on Otto and Cummiskey (1990), as in the past. Other relevant data sources,
including assumed population and fishery parameters, are presented in Appendix A, which also provides a detailed
description of the model configuration used for this assessment.

# E. Analytic Approach

## History of Modeling Approaches for this Stock

A four-stage catch-survey-analysis (CSA) assessment model was used before 2011 to estimate abundance and biomass and
prescribe fishery quotas for the SMBKC stock. The four-stage CSA is similar to a full
length-based analysis, the major difference being coarser length groups, which are more suited to a small stock with
consistently low survey catches. In this approach, the abundance of male crab with a CL $\ge$ 90 mm is modeled in
terms of four crab stages: stage 1: 90-104 mm CL; stage 2: 105-119 mm CL; stage 3: newshell 120-133 mm CL; and stage 4:
oldshell $\ge$ 120 mm CL and newshell $\ge$ 134 mm CL. Motivation for these stage definitions comes from the fact that
for management of the SMBKC stock, male crab measuring $\ge$ 105 mm CL are considered mature, whereas 120 mm CL is
considered a proxy for the legal size of 5.5 in carapace width, including spines. Additional motivation for these stage
definitions comes from an estimated average growth increment of about 14 mm per molt for SMBKC (Otto and Cummiskey
1990).

Concerns about the pre-2011 assessment model led to the CPT and SSC recommendations that included development of an
alternative model with provisional assessment based on survey biomass or some other index of abundance. An alternative
3-stage model was proposed to the CPT in May 2011, but a survey-based approach was requested for the Fall
2011 assessment. In May 2012 the CPT approved a slightly revised and better documented version of the alternative model
for assessment. Subsequently, the model developed and used since  2012 was a variant of the previous four-stage SMBKC
CSA model and similar in complexity to that described by Collie et al. (2005). Like the earlier model, it considered
only male crab $\ge$ 90 mm in CL, but combined stages 3 and 4 of the earlier model, resulting in three stages
(male size classes) defined by CL measurements of (1) 90-104 mm, (2) 105-119 mm, and (3) 120 mm+ (i.e., 120 mm and
above). This consolidation was driven by concern about the accuracy and consistency of shell-condition information,
which had been used in distinguishing stages 3 and 4 of the earlier model.

In 2016 the accepted SMBKC assessment model made use of the modeling framework Gmacs (Webber et al. 2016). In that
assessment,  an effort was made to match the 2015 SMBKC stock assessment model to bridge a framework which 
provided greater flexibility and opportunity to evaluate model assumptions more fully.

## Assessment Methodology

This assessment model again uses the modeling framework Gmacs and is detailed in Appendix A.

## Model Selection and Evaluation

Five models were presented in the previous assessment. This year, four models are presented with the  reference model
being the same configuration as approved last year (Ianelli et al. 2017), two sensitivities are considered, one with a  different treatment of NMFS
bottom trawl survey (BTS) data using a geo-spatial model (VAST; Thorson and Barnett 2017; Appendix C). A second 
sensitivity was constructed which weights the
survey data more heavily. In addition to these sensitivities, we evaluated the impacts of adding new data 
to the reference model. In summary, the following lists the models presented and the naming convention  used:

1. **2017 Model**: the 2017 recommended model without any new data

2. **BTS**: adds in the 2018 bottom trawl survey (BTS) data

3. **BTS and pot**: as with previous but including the 2018 ADFG pot survey data (Model 16.0 or "reference case")

3. **VAST**: applies a geo-spatial delta-GLMM model (Thorson and Barnett 2017) to the BTS data which provides a different BTS index. See appendix B for 
  details and diagnostics. This is a preliminary examination as more work is needed to ensure options for the BTS CPUE data
   were specified appropriately.

3. **Fit survey**: an exploratory scenario that's the same as the reference model except the NMFS trawl survey is up-weighted by
   $\lambda^\text{NMFS}=$ `r M[[5]]$cpue_lambda[1]` and the ADF&G pot survey is 
   up-weighted by $\lambda^\text{ADFG}=$ `r M[[5]]$cpue_lambda[2]`. 

Note that SSC convention would label these (item 3 above) as model 16.0 (the
model first developed in that year). Since only a few models are presented here, for simplicity
we labeled model 16.0 as "reference" and for the others, we used the  simple naming convention
presented above.

## Results
### a. Sensitivity to new data
Results for scenarios are provided with comparisons to the 2017 model and sensitivity new data are shown in Figures
\ref{fig:surv1} and \ref{fig:surv2} with recruitment and spawning biomass shown in Figures \ref{fig:rec1} and \ref{fig:ssb1}, respectively. 
 The fits to survey CPUEs and spawning biomass show that the addition of new data results in more of a decline than
 in the 2017 assessment, especially with the addition of the pot survey. 

### b. Alternative NMFS bottom-trawl survey index
Results comparing model fits between the VAST model and the reference case show different time-series
of data and a different model fit (Figure \ref{fig:surv3}). The effect on spawning biomass suggests estimates were consistently
higher since 1990 compared to the reference model (Figure  \ref{fig:ssb2}).

### c. Effective sample sizes and weighting factors
Observed and estimated effective sample sizes are compared in Table \ref{tab:effn}. 
Data weighting factors, standard deviation of normalized residuals (SDNRs), and median absolute residual (MAR) are 
presented in Table \ref{tab:data_weighting}. The SDNR for the trawl survey
is acceptable at 1.66 in the reference model. Francis (2011) weighting was applied in 2017
but given the relatively few size
bins in this assessment, this application was suspended this year.  
The SDNRs for the pot surveys show a similar pattern in each of the
scenarios, but are much higher suggesting an inconsistency between 
the pot survey data and the model structure and other data
components. Rather than re-weighting, we chose to retain the values 
as specified, noting that down-weighting these data would effectively 
exclude the signal from this series. 
The MAR values for the trawl and pot surveys shows the same pattern among each of the
scenarios as the SDNR. The SDNR and MAR values for the trawl survey and pot survey size compositions were relatively good,
ranging from 0.54 to 0.73 for the reference case.
The SDNRs for the directed pot fishery and other size compositions were  similar to previous estimates.

### d. Parameter estimates
Model parameter estimates for each of the Gmacs scenarios are summarized in Tables \ref{tab:est_pars_ref},
\ref{tab:est_pars_vast}, and \ref{tab:est_pars_all}. These parameter
estimates are compared in Table \ref{tab:est_pars_all}. Negative log-likelihood values and management measures for each
of the model configurations are compared in Tables \ref{tab:management_quants} through \ref{tab:likelihood_components}.

There are some differences in parameter estimates among models as 
reflected in the log-likelihood components and the management quantities. 
The parameter estimates in the "fit survey" scenario differ the most, as expected, particularly the estimate of the ADF&G pot survey
catchability ($q$) (see Table \ref{tab:est_pars_all}). Also, the residuals for recruitment in the first size group are large for these
model runs, presumably because higher estimates of recruits in some years are required by the model
 to match the observed biomass trends. 

Selectivity estimates show some variability between models (Figure \ref{fig:selectivity}). Estimated recruitment is
variable over time for all models and in recent years is well below average (Figure \ref{fig:recruitment}).  Estimated
mature male biomass on 15 February also fluctuates considerably (Figure \ref{fig:mmb}). Estimated natural mortality each
year ($M_t$) is presented in Figure \ref{fig:M_t}.

### e. Evaluation of the fit to the data.
The model fits to total male ($\ge$ 90 mm CL) trawl survey biomass tend to miss the recent peak around 2010 and is
slightly above the 2017 value for the key sensitivities (Figures \ref{fig:trawl_survey_biomass}).  All of the models fit
the pot survey CPUE poorly (Figure \ref{fig:pot_survey_cpue}. For both surveys the standardized residuals tend to have
similar patterns with some improvement (generally) for the  VAST model (Figures \ref{fig:bts_resid_nmfs} and
\ref{fig:bts_resid_adfg}).

Fits to the size compositions for trawl survey, pot survey, and commercial observer data are reasonable but miss the largest
size category in some years (Figures \ref{fig:sc_pot}, \ref{fig:sc_pot_discarded}, and \ref{fig:sc_trawl_discarded}) 
for all scenarios. Representative
residual  plots of the composition data fits are generally poor (Figures \ref{fig:sc_res_ref} and \ref{fig:sc_res_fit_survey}).
The model fits to different types of retained and discarded catch values performed as expected given the assumed levels
of uncertainty on the input data (Figure  \ref{fig:fit_to_catch} ).

Unsurprisingly, the **Fit surveys** model fits the 
the NMFS survey biomass and ADF&G pot survey CPUE  data better but still has a similar residual pattern
(Figures \ref{fig:trawl_survey_biomass} and \ref{fig:pot_survey_cpue}). 
It is worth noting that that this scenario (included for exploratory purposes) resulted in worse SDNR and MAR values for the two abundance indices.

### f. Retrospective and historical analyses
This is only the second year a formal assessment model developed for this stock. As such, retrospective patterns and 
historical analyses relative to fisheries impacts are limited.

### g. Uncertainty and sensitivity analyses.
Estimated standard deviations of parameters and selected management measures for the models are summarized in
Tables \ref{tab:est_pars_ref}, \ref{tab:est_pars_vast}, and \ref{tab:est_pars_fit_survey} 
(compiled in Table \ref{tab:est_pars_all}). Probabilities for mature male biomass and OFL in 2017 are
presented in Section F.


### h. Comparison of alternative model scenarios.
The estimates of mature male biomass (Figure \ref{fig:mmb}), for the **Fit survey** sensitivity 
differs from the other models due to a low value for pot survey catchability being estimated (which tends to scale the population
estimate). This existng scenario results in a lower MMB from the mid-1980s
through to the late-1990s, and is again lower in the most recent 5 years. This scenario upweights both the trawl 
and pot surveys abundance indices  and represents a model run
that places greater emphasis on the abundance indices.

In summary, the use of the reference model for management purposes is preferred since it provides the best fit to the
data and is consistent with previous model specifications.  Research on alternative model specifications (e.g., natural
mortality variability) was limited this year. The VAST model may take better account of spatial
processes but requires more research to ensure it has been appropriately applied and the assumptions are reasonable.
Consequently, the reference model appears reasonable and appropriate for ACL and OFL  determinations for this stock in
2017. Nonetheless, the **Fit surveys** model, while difficult to statistically justify, portends a more dire stock status
(see below) and should highlight the caution needed in managing this resource.


# F. Calculation of the OFL and ABC
The overfishing level (OFL) is the fishery-related mortality biomass associated with fishing mortality $F_\mathit{OFL}$. The SMBKC stock is currently managed as Tier 4, and only a Tier 4 analysis is presented here. Thus given stock estimates or suitable proxy values of $B_\mathit{MSY}$ and $F_\mathit{MSY}$, along with two additional parameters $\alpha$ and $\beta$, $F_\mathit{OFL}$ is determined by the control rule
\begin{align}
    F_\mathit{OFL} &= 
    \begin{cases}
        F_\mathit{MSY}, &\text{ when } B/B_\mathit{MSY} > 1\\
        F_\mathit{MSY} \frac{\left( B/B_\mathit{MSY} - \alpha \right)}{(1 - \alpha)}, &\text{ when } \beta < B/B_\mathit{MSY} \le 1
    \end{cases}\\
    F_\mathit{OFL} &< F_\mathit{MSY} \text{ with directed fishery } F = 0 \text{ when } B/B_\mathit{MSY} \le \beta \notag
\end{align}
where $B$ is quantified as mature-male biomass (MMB) at mating with time of mating assigned a nominal date of 15
February. Note that as $B$ itself is a function of the fishing mortality $F_\mathit{OFL}$ (therefore numerical
approximation of $F_\mathit{OFL}$ is required). As implemented for this assessment, all calculations proceed according
to the model equations given in Appendix A. $F_\mathit{OFL}$ is taken to be full-selection fishing mortality in the
directed pot fishery and groundfish trawl and fixed-gear fishing mortalities set at their model geometric mean values
over years for which there are data-based estimates of bycatch-mortality biomass.

The currently recommended Tier 4 convention is to use the full assessment period, currently `r M[[rec_mod]]$spr_syr`-
`r M[[rec_mod]]$spr_nyr`, to define a $B_\mathit{MSY}$ proxy in terms of average estimated MMB and to set $\gamma$ = 1.0 with
assumed stock natural mortality $M$ = 0.18 $\text{yr}^{-1}$ in setting the $F_\mathit{MSY}$ proxy value $\gamma M$. The
parameters $\alpha$ and $\beta$ are assigned their default values $\alpha$ = 0.10 and $\beta$ = 0.25. The
$F_\mathit{OFL}$, OFL, ABC, and MMB in 2018 for all scenarios are summarized in Table \ref{tab:management_quants}. The ABC
is 80% of the OFL.
```{r management_quants, results = "asis"}
df <- NULL
for (ii in mod_scen)
{
    x      <- M[[ii]]
    mmb    <- x$ssb[length(x$ssb)]; names(mmb) <- paste0("$\\text{MMB}_{", x$mod_yrs[length(x$mod_yrs)], "}$")
    fofl   <- x$spr_fofl; names(fofl)          <- "$F_\\text{OFL}$"
    OFL    <- x$spr_cofl; names(OFL)           <- paste0("$\\text{OFL}_{", x$mod_yrs[length(x$mod_yrs)], "}$")
    Bmsy   <- x$spr_bspr; names(Bmsy)          <- "$B_\\text{MSY}$"
    B_Bmsy <- mmb/Bmsy; names(B_Bmsy)          <- "$MMB/B_\\text{MSY}$"
    ABC    <- OFL * 0.8; names(ABC)            <- paste0("$\\text{ABC}_{", x$mod_yrs[length(x$mod_yrs)], "}$")
    v      <- c(mmb, Bmsy, fofl, OFL, ABC)
    df     <- cbind(df, v)
}
df <- data.frame(rownames(df), df, row.names = NULL)
names(df) <- c("Component",mod_names[mod_scen])
tab <- xtable(df, caption = "Comparisons of management measures for the model scenarios. Biomass and OFL are in tons.", label = "tab:management_quants", digits = 3)
print(tab, caption.placement = "top", include.rownames = FALSE, sanitize.text.function = function(x){x})
```


# G. Rebuilding Analysis

This stock is not currently subject to a rebuilding plan. However, interpretation of the point estimate for the reference case
suggests that the mature male biomass is below 50% of $B_\mathit{MSY}$ but slightly above for the "VAST" model configuration (Table \ref{tab:management_quants} ).

# H. Data Gaps and Research Priorities

The following topics have been listed as areas where more research on SMBKC is needed:

  1. Growth increments and molting probabilities as a function of size.

  2. Trawl survey catchability and selectivities.

  3. Temporal changes in spatial distributions near the island.

  4. Natural mortality.

# I. Projections and outlook

The outlook for recruitment is pessimistic and the abundance relative to the proxy $B_\mathit{MSY}$ is low. 
The NMFS survey results in 2018 noted ocean conditions warmer than 
normal with an absence of a "cold pool" in the region. This could have detrimental effects on the SMBKC stocks and
should be carefully monitored. Relative to the impact of historical fishing, we again conducted a
 "dynamic-$B_0$" analysis. This procedure simply projects
the population based on estimated recruitment but removes the effect of fishing. For the reference case,
this suggests that the impact of fishing has reduced to stock to about
  `r .get_dynB0_df(Mbase) %>% transmute(round(100*ssb,0)) %>% slice(n())`% of what it would have been in the absence of
  fishing (Figure \ref{fig:dynB0}) . The other non-fishing contributors to the observed depleted stock trend (ignoring stock-recruit
  relationship) may reflect variable  survival rates due to environmental conditions and also range shifts. 

# J. Acknowledgements

We thank the Crab Plan Team and AFSC staff for reviewing an earlier draft of this report 
and Andre Punt for his input into refinements to the Gmacs model code.

\pagebreak 

## Reference
Alaska Department of Fish and Game (ADF&G). 2013. Crab observer training and deployment manual. Alaska Department of Fish and Game Shellfish Observer Program, Dutch Harbor. Unpublished.

Collie, J.S., A.K. Delong, and G.H. Kruse. 2005. Three-stage catch-survey analysis applied to blue king crabs. Pages 683-714 [In] Fisheries assessment and management in data-limited situations. University of Alaska Fairbanks, Alaska Sea Grant Report 05-02, Fairbanks.

Daly, B., R. Foy, and C. Armistead. 2014. The 2013 eastern Bering Sea continental shelf bottom trawl survey: results for commercial crab species. NOAA Technical Memorandum 295, NMFS-AFSC.

Donaldson, W.E., and S.C. Byersdorfer. 2005. Biological field techniques for lithodid crabs. University of Alaska Fairbanks, Alaska Sea Grant Report 05-03, Fairbanks.

Fitch, H., M. Deiman, J. Shaishnikoff, and K. Herring. 2012. Annual management report for the commercial and subsistence shellfish fisheries of the Bering Sea, 2010/11. Pages 75-1776 [In] Fitch, H., M. Schwenzfeier, B. Baechler, T. Hartill, M. Salmon, and M. Deiman, E.

Evans, E. Henry, L. Wald, J. Shaishnikoff, K. Herring, and J. Wilson. 2012. Annual management report for the commercial and subsistence shellfish fisheries of the Aleutian Islands, Bering Sea and the Westward Region’s Shellfish Observer Program, 2010/11. Alaska Department of Fish and Game, Fishery Management Report No. 12-22, Anchorage.

Fournier, D.A., H.J. Skaug, J. Ancheta, J. Ianelli, A. Magnusson, M.N. Maunder, A. Nielsen, and J. Sibert. 2012. AD Model Builder: using automatic differentiation for statistical inference of highly parameterized complex nonlinear models. Optim. Methods Softw. 27:233-249.

Francis, R.I.C.C. 2011. Data weighting in statistical fisheries stock assessment models. Can. J. Fish. Aquat. Sci. 68: 1124-1138.

Gaeuman, W.B. 2013. Summary of the 2012/13 mandatory crab observer program database for the Bering Sea/Aleutian Islands commercial crab fisheries. Alaska Department of Fish and Game, Fishery Data Series No. 13-54, Anchorage. 

Gish, R.K., V.A. Vanek, and D. Pengilly. 2012. Results of the 2010 triennial St. Matthew Island blue king crab pot survey and 2010/11 tagging study. Alaska Department of Fish and Game, Fishery Management Report No. 12-24, Anchorage.

Ianelli, J., D. Webber, and J. Zheng, 2017. Stock assessment of Saint Matthews Island blue king crab. North Pacific Fishery Management Council. Anchorage AK.

Jensen, G.C., and D.A. Armstrong. 1989. Biennial reproductive cycle of blue king crab, *Paralithodes platypus*, at the Pribilof Islands, Alaska and comparison to a congener, *P. camtschatica*. Can. J. Fish. Aquat. Sci. 46: 932-940.

Moore, H., L.C. Byrne, and D. Connolly. 2000. Alaska Department of Fish and Game summary of the 1998 mandatory shellfish observer program database. Alaska Dept. Fish and Game, Commercial Fisheries Division, Reg. Inf. Rep. 4J00-21, Kodiak.

North Pacific Fishery Management Council (NPFMC). 1998. Fishery Management Plan for Bering Sea/Aleutian Islands king and Tanner crabs. North Pacific Fishery Management Council, Anchorage.

North Pacific Fishery Management Council (NPFMC). 1999. Environmental assessment/regulatory impact review/initial regulatory flexibility analysis for Amendment 11 to the Fishery Management Plan for Bering Sea/Aleutian Islands king and Tanner crabs. North Pacific Fishery Management Council, Anchorage.

North Pacific Fishery Management Council (NPFMC). 2000. Environmental assessment/regulatory impact review/initial regulatory flexibility analysis for proposed Amendment 15 to the Fishery Management Plan for king and Tanner crab fisheries in the Bering Sea/Aleutian Islands and regulatory amendment to the Fishery Management Plan for the groundfish fishery of the Bering Sea and Aleutian Islands area: A rebuilding plan for the St. Matthew blue king crab stock. North Pacific Fishery Management Council, Anchorage. Draft report.

North Pacific Fishery Management Council (NPFMC). 2007. Public Review Draft: Environmental assessment for proposed Amendment 24 to the Fishery Management Plan for Bering Sea and Aleutian Islands king and Tanner crabs to revise overfishing definitions. 14 November 2007. North Pacific Fishery Management Council, Anchorage.

Otto, R.S. 1990. An overview of eastern Bering Sea king and Tanner crab fisheries. Pages 9-26 [In] Proceedings of the international symposium on king and Tanner crabs. University of Alaska Fairbanks, Alaska Sea Grant Program Report 90-4, Fairbanks.

Otto, R.S., and P.A. Cummiskey. 1990. Growth of adult male blue king crab (*Paralithodes platypus*). Pages 245-258 [In] Proceedings of the international symposium on king and Tanner crabs. University of Alaska Fairbanks, Alaska Sea Grant Report 90-4, Fairbanks.

Paul, J.M., A. J. Paul, R.S. Otto, and R.A. MacIntosh. 1991. Spermatophore presence in relation to carapace length for eastern Bering Sea blue king crab (*Paralithodes platypus*, Brandt, 1850) and red king crab (*P. camtschaticus*, Tilesius, 1815). J. Shellfish Res. 10: 157-163.

Pengilly, D. and D. Schmidt. 1995. Harvest Strategy for Kodiak and Bristol Bay red king crab and St. Matthew Island and Pribilof blue king crab. Alaska Department of Fish and Game, Commercial Fisheries Management and Development Division, Special Publication Number 7, Juneau.

Somerton, D.A., and R.A. MacIntosh. 1983. The size at sexual maturity of blue king crab, Paralithodes platypus, in Alaska. Fishery Bulletin 81: 621-828.

Thorson, J.T., and L.A.K. Barnett. 2017. Comparing estimates of abundance trends and distribution shifts using single- and multispecies models of fishes and biogenic habitat. ICES Journal of Marine Science 75:1311-1321.

Thorson, J.T., J.N. Ianelli, E. Larsen, L. Ries, M.D. Scheuerell, C. Szuwalski, and E. Zipkin. 2016. Joint dynamic species distribution models: a tool for community ordination and spatiotemporal monitoring. Glob. Ecol. Biogeogr. 25(9): 1144–1158. geb.12464. 

Thorson, J.T., Scheuerell, M.D., Shelton, A.O., See, K.E., Skaug, H.J., and Kristensen, K. 2015. Spatial factor analysis: a new tool for estimating joint species distributions and correlations in species range. Methods Ecol. Evol. 6(6): 627–637. doi:10.1111/2041-210X.12359. 

Webber, D., J. Zheng, and J. Ianelli, 2016. Stock assessment of Saint Matthews Island Blue King Crab. North Pacific Fishery Managment Council. Anchorage AK.

Zheng, J. 2005. A review of natural mortality estimation for crab stocks: data-limited for every stock? Pages 595-612 [In] Fisheries Assessment and Management in Data-Limited Situations. University of Alaska Fairbanks, Alaska Sea Grant Program Report 05-02, Fairbanks.

Zheng, J., and G.H. Kruse. 2002. Assessment and management of crab stocks under uncertainty of massive die-offs and rapid changes in survey catchability. Pages 367-384 [In] A.J. Paul,E.G. Dawe, R. Elner, G.S. Jamieson, G.H. Kruse, R.S. Otto, B. Sainte-Marie, T.C. Shirley, and D. Woodby (eds.). Crabs in Cold Water Regions: Biology, Management, and Economics. University of Alaska Fairbanks, Alaska Sea Grant Report 02-01, Fairbanks.

Zheng, J., M.C. Murphy, and G.H. Kruse. 1997. Application of catch-survey analysis to blue king crab stocks near Pribilof and St. Matthew Islands. Alaska Fish. Res. Bull. 4:62-74.


\pagebreak 

## Tables 



\pagebreak 

## Figures


\newpage 


